---
title: "Soccer Ratings Backtest"
subtitle: "Predictions vs Actuals — FanTeam Monster"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
    toc_depth: 3
    number_sections: false
    code_folding: hide
    fig_width: 10
    fig_height: 6
    df_print: kable
params:
  gameweeks: NULL
  force_refresh: false
---

```{css, echo=FALSE}
/* ---- Custom Theme ---- */
:root {
  --sage: #A3BE8C;
  --sage-dark: #8FAF78;
  --coral: #D08770;
  --coral-dark: #BF7460;
  --primary: #3B3226;
  --bg: #FAF8F5;
  --muted: #8C8278;
}

body {
  background-color: var(--bg);
  color: var(--primary);
  font-family: 'Source Sans Pro', 'Segoe UI', sans-serif;
}

h1, h2, h3, h4 { color: var(--primary); font-weight: 700; }

h1.title {
  font-size: 2.2em;
  border-bottom: 4px solid var(--sage);
  padding-bottom: 0.3em;
}

h3 { border-left: 4px solid var(--sage); padding-left: 0.6em; }

/* Metric cards */
.metric-row { display: flex; gap: 1em; flex-wrap: wrap; margin: 1em 0; }
.metric-card {
  background: white;
  border: 2px solid #e8e4df;
  border-radius: 8px;
  padding: 1em 1.5em;
  flex: 1;
  min-width: 150px;
  text-align: center;
}
.metric-card .value {
  font-size: 2em;
  font-weight: 800;
  color: var(--primary);
  line-height: 1.2;
}
.metric-card .label {
  font-size: 0.85em;
  color: var(--muted);
  text-transform: uppercase;
  letter-spacing: 0.05em;
}
.metric-card.good .value { color: var(--sage-dark); }
.metric-card.bad .value { color: var(--coral-dark); }
.metric-card.neutral .value { color: var(--primary); }

/* Rating badges */
.badge-chalk   { background: #D08770; color: white; padding: 2px 8px; border-radius: 4px; font-weight: 600; font-size: 0.85em; }
.badge-popular { background: #EBCB8B; color: #3B3226; padding: 2px 8px; border-radius: 4px; font-weight: 600; font-size: 0.85em; }
.badge-mod     { background: #A3BE8C; color: white; padding: 2px 8px; border-radius: 4px; font-weight: 600; font-size: 0.85em; }
.badge-low     { background: #B8C9A3; color: white; padding: 2px 8px; border-radius: 4px; font-weight: 600; font-size: 0.85em; }
.badge-ghost   { background: #D8D8D8; color: #666; padding: 2px 8px; border-radius: 4px; font-weight: 600; font-size: 0.85em; }

/* Table styling */
table { width: 100%; border-collapse: collapse; }
th { background: var(--primary); color: white; padding: 8px 12px; text-align: left; font-size: 0.85em; }
td { padding: 6px 12px; border-bottom: 1px solid #e8e4df; font-size: 0.9em; }
tr:hover td { background: #f0ede8; }

/* Interpretation boxes */
.interpretation {
  background: white;
  border-left: 4px solid var(--sage);
  padding: 0.8em 1.2em;
  margin: 1em 0;
  font-size: 0.95em;
  color: #555;
}
.interpretation strong { color: var(--primary); }
.interpretation.warning { border-left-color: var(--coral); }

/* Section breaks */
hr { border: none; border-top: 2px solid #e8e4df; margin: 2em 0; }
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  fig.align = "center",
  out.width = "100%"
)

# ---- Load app infrastructure ----
# Source from project root (assumes knit directory is project root)
suppressPackageStartupMessages({
  source("global.R", local = TRUE)
  source("R/soccer/soccer_ratings.R", local = TRUE)
})

library(knitr)
library(scales)

# ---- Colour palette matching app ----
pal <- list(
  sage      = "#A3BE8C",
  sage_dark = "#8FAF78",
  sage_light = "#C5D9B2",
  coral      = "#D08770",
  coral_dark = "#BF7460",
  coral_light = "#E8B4A2",
  primary    = "#3B3226",
  muted      = "#8C8278",
  bg         = "#FAF8F5",
  yellow     = "#EBCB8B",
  blue       = "#5E81AC",
  grid       = "#E8E4DF"
)

# Rating tier colour maps
cash_colors <- c(
  "Cash Lock"    = pal$sage_dark,
  "Strong Play"  = pal$sage,
  "Decent"       = pal$yellow,
  "Risky"        = pal$coral_light,
  "Hard Pass"    = pal$coral_dark
)

gpp_colors <- c(
  "Smash Play"   = pal$sage_dark,
  "Strong GPP"   = pal$sage,
  "Playable"     = pal$yellow,
  "Meh"          = pal$coral_light,
  "Fade"         = pal$coral_dark
)

pos_colors <- c(
  GK  = pal$yellow,
  DEF = pal$blue,
  MID = pal$sage,
  FWD = pal$coral
)

# ---- ggplot theme ----
theme_backtest <- function(base_size = 12) {
  theme_minimal(base_size = base_size) +
    theme(
      text = element_text(color = pal$primary),
      plot.title = element_text(face = "bold", size = base_size * 1.2, margin = margin(b = 8)),
      plot.subtitle = element_text(color = pal$muted, size = base_size * 0.9, margin = margin(b = 12)),
      panel.grid.minor = element_blank(),
      panel.grid.major = element_line(color = pal$grid, linewidth = 0.4),
      axis.title = element_text(color = pal$muted, size = base_size * 0.9),
      axis.text = element_text(color = pal$primary),
      legend.position = "top",
      legend.text = element_text(size = base_size * 0.85),
      strip.text = element_text(face = "bold", size = base_size * 0.95),
      plot.background = element_rect(fill = pal$bg, color = NA),
      panel.background = element_rect(fill = "white", color = NA)
    )
}

# ---- Helper: styled kable ----
styled_table <- function(df, caption = NULL, digits = 2) {
  kable(df, digits = digits, format = "html", caption = caption, escape = FALSE,
        table.attr = 'class="table table-striped table-hover table-condensed"')
}

# ---- Helper: metric card HTML ----
metric_card <- function(value, label, class = "neutral", fmt = NULL) {
  if (!is.null(fmt)) value <- sprintf(fmt, value)
  sprintf('<div class="metric-card %s"><div class="value">%s</div><div class="label">%s</div></div>',
          class, value, label)
}
```

```{r load-data, cache=FALSE}
# ---- Discover available gameweeks ----
available_gws <- get_fanteam_soccer_gameweeks()
test_gws <- if (!is.null(params$gameweeks)) {
  intersect(params$gameweeks, available_gws)
} else {
  available_gws
}

# ---- Load shared data ----
stats_overview <- tryCatch(load_fanteam_stats_overview(force_refresh = params$force_refresh),
                           error = function(e) NULL)
stats_detail   <- tryCatch(load_fanteam_stats_detail(force_refresh = params$force_refresh),
                           error = function(e) NULL)
own_history    <- tryCatch(load_ownership_history(),
                           error = function(e) NULL)

# Identify columns in detail data
pts_col <- NULL; gw_col <- NULL
if (!is.null(stats_detail) && nrow(stats_detail) > 0) {
  pts_col <- intersect(c("pts", "total_pts"), names(stats_detail))[1]
  gw_col  <- intersect(c("gw", "gameweek"), names(stats_detail))[1]
  if (!is.null(pts_col)) stats_detail[[pts_col]] <- as.numeric(stats_detail[[pts_col]])
  if (!is.null(gw_col))  stats_detail[[gw_col]]  <- as.integer(stats_detail[[gw_col]])
}

# Available GWs in detail (for actual points)
detail_gws <- if (!is.null(gw_col)) sort(unique(stats_detail[[gw_col]])) else integer(0)
```

```{r process-gameweeks, cache=FALSE}
# ---- Process each gameweek through the ratings pipeline ----
process_gw <- function(gw) {
  salaries <- tryCatch(load_fanteam_soccer_with_logos(gw), error = function(e) NULL)
  if (is.null(salaries) || nrow(salaries) == 0) return(NULL)
  
  odds <- tryCatch(load_fanteam_odds(gw), error = function(e) NULL)
  
  # Build matchup context
  matchup_ctx <- NULL
  if (!is.null(odds) && nrow(odds) > 0) {
    slate_teams <- unique(salaries$team_normalized[!is.na(salaries$team_normalized)])
    matchup_ctx <- tryCatch({
      odds %>%
        filter(odds_team_normalized %in% slate_teams) %>%
        transmute(
          team_normalized = odds_team_normalized,
          opponent = odds_opponent_normalized,
          home_away = home_away,
          win_pct = as.numeric(win_pct),
          implied_goals = as.numeric(implied_team_goals),
          implied_opp_goals = as.numeric(implied_opp_goals),
          implied_total = implied_goals + implied_opp_goals,
          cs_prob = round(exp(-implied_opp_goals) * 100, 1)
        )
    }, error = function(e) NULL)
  }
  
  pool <- salaries %>% mutate(pos = position)
  
  # Merge stats
  if (!is.null(stats_overview) && nrow(stats_overview) > 0) {
    stats_cols <- stats_overview %>%
      select(any_of(c("name", "team", "pos", "x1_mp", "total_pts", "avg_pts", "picked_by")))
    
    if (!is.null(stats_detail) && !is.null(pts_col) && nrow(stats_detail) > 0) {
      player_perf <- stats_detail %>%
        group_by(name) %>%
        summarise(
          n_games = n(),
          ppg_hist = mean(.data[[pts_col]], na.rm = TRUE),
          pts_floor = { p <- sort(.data[[pts_col]][!is.na(.data[[pts_col]])]); if (length(p) >= 4) p[2] else if (length(p) > 0) p[1] else NA_real_ },
          pts_ceiling = { p <- sort(.data[[pts_col]][!is.na(.data[[pts_col]])], decreasing = TRUE); if (length(p) >= 4) p[2] else if (length(p) > 0) p[1] else NA_real_ },
          downside_dev = { p <- .data[[pts_col]][!is.na(.data[[pts_col]])]; a <- mean(p, na.rm = TRUE); b <- p[p < a]; if (length(b) > 1) sqrt(mean((b - a)^2)) else NA_real_ },
          .groups = "drop"
        ) %>%
        mutate(sortino_raw = if_else(!is.na(ppg_hist) & !is.na(downside_dev) & downside_dev > 0, ppg_hist / downside_dev, NA_real_))
      
      k_prior <- 10
      pos_avgs <- player_perf %>%
        left_join(stats_cols %>% select(name, pos) %>% distinct(), by = "name") %>%
        filter(n_games >= 6, !is.na(sortino_raw)) %>%
        group_by(pos) %>%
        summarise(pos_avg_sort = mean(sortino_raw, na.rm = TRUE), .groups = "drop")
      
      player_perf <- player_perf %>%
        left_join(stats_cols %>% select(name, pos) %>% distinct(), by = "name") %>%
        left_join(pos_avgs, by = "pos") %>%
        mutate(
          pos_avg_sort = coalesce(pos_avg_sort, mean(sortino_raw, na.rm = TRUE)),
          sortino = if_else(!is.na(sortino_raw) & !is.na(n_games),
                             (n_games * sortino_raw + k_prior * pos_avg_sort) / (n_games + k_prior), NA_real_)
        ) %>%
        select(name, n_games, ppg_hist, pts_floor, pts_ceiling, sortino)
      
      stats_cols <- stats_cols %>% left_join(player_perf, by = "name")
    }
    
    pool <- pool %>%
      mutate(match_name = tolower(trimws(player))) %>%
      left_join(stats_cols %>% mutate(match_name = tolower(trimws(name))),
                by = "match_name", suffix = c("", "_stats")) %>%
      select(-match_name) %>%
      mutate(
        pos = coalesce(pos_stats, pos),
        ppg = coalesce(as.numeric(ppg_hist), as.numeric(avg_pts), 0),
        pts_floor = coalesce(pts_floor, 0),
        pts_ceiling = coalesce(pts_ceiling, 0),
        sortino = coalesce(sortino, 0),
        n_games = coalesce(n_games, 0L),
        form = ppg
      ) %>%
      select(-any_of(c("pos_stats", "name_stats", "team_stats")))
  } else {
    pool <- pool %>% mutate(ppg = 0, pts_floor = 0, pts_ceiling = 0, sortino = 0,
                             n_games = 0L, form = 0, picked_by = NA_real_)
  }
  
  if (!is.null(matchup_ctx)) {
    pool <- pool %>% left_join(matchup_ctx %>% select(team_normalized, opponent, home_away), by = "team_normalized")
  } else {
    pool <- pool %>% mutate(opponent = NA_character_, home_away = NA_character_)
  }
  
  pool <- calculate_matchup_quality(pool, matchup_ctx)
  pool <- estimate_ownership(pool, historical = own_history)
  pool <- rate_players(pool)
  pool <- pool %>% mutate(value = round(ppg / pmax(salary, 3), 2))
  
  # Join actual points
  pool$actual_pts <- NA_real_
  if (!is.null(stats_detail) && !is.null(pts_col) && !is.null(gw_col) && gw %in% detail_gws) {
    gw_actuals <- stats_detail %>%
      filter(.data[[gw_col]] == gw) %>%
      transmute(match_name = tolower(trimws(name)), actual_pts = as.numeric(.data[[pts_col]]))
    if (nrow(gw_actuals) > 0) {
      pool <- pool %>%
        mutate(match_name = tolower(trimws(player))) %>%
        left_join(gw_actuals, by = "match_name") %>%
        mutate(actual_pts = coalesce(actual_pts.y, actual_pts.x)) %>%
        select(-any_of(c("actual_pts.x", "actual_pts.y", "match_name")))
    }
  }
  
  # Join actual ownership
  pool$actual_own_pct <- NA_real_; pool$actual_captain_pct <- NA_real_
  if (!is.null(own_history) && nrow(own_history) > 0) {
    gw_own <- own_history %>%
      filter(gameweek == gw, !is.na(own_pct)) %>%
      transmute(match_name = tolower(trimws(player)), actual_own_pct = own_pct,
                actual_captain_pct = coalesce(captain_pct, NA_real_))
    if (nrow(gw_own) > 0) {
      pool <- pool %>%
        mutate(match_name = tolower(trimws(player))) %>%
        left_join(gw_own, by = "match_name") %>%
        mutate(actual_own_pct = coalesce(actual_own_pct.y, actual_own_pct.x),
               actual_captain_pct = coalesce(actual_captain_pct.y, actual_captain_pct.x)) %>%
        select(-any_of(c("actual_own_pct.x", "actual_own_pct.y",
                          "actual_captain_pct.x", "actual_captain_pct.y", "match_name")))
    }
  }
  
  pool$gameweek <- gw
  return(pool)
}

# Run all gameweeks
all_results <- lapply(sort(test_gws, decreasing = TRUE), function(gw) {
  tryCatch(process_gw(gw), error = function(e) NULL)
})
all_results <- Filter(Negate(is.null), all_results)
data <- bind_rows(all_results)

has_actuals <- sum(!is.na(data$actual_pts)) > 0
has_own_actuals <- sum(!is.na(data$actual_own_pct)) > 0
gws_with_pts <- sort(unique(data$gameweek[!is.na(data$actual_pts)]))
n_player_gws <- sum(!is.na(data$actual_pts))
```

## Overview {.tabset}

```{r overview-cards, results='asis'}
cat('<div class="metric-row">')
cat(metric_card(length(test_gws), "Gameweeks<br>with salaries", "neutral"))
cat(metric_card(length(gws_with_pts), "Gameweeks<br>with actuals", if (length(gws_with_pts) > 0) "good" else "bad"))
cat(metric_card(comma(nrow(data)), "Player-GW<br>records", "neutral"))
cat(metric_card(comma(n_player_gws), "With actual<br>points", if (n_player_gws > 0) "good" else "bad"))
cat('</div>')

if (length(gws_with_pts) > 0) {
  cat(sprintf('<div class="interpretation">Actuals available for GW: <strong>%s</strong>. Stats overview has <strong>%s</strong> players. Detail data covers <strong>%d</strong> gameweeks.</div>',
              paste(gws_with_pts, collapse = ", "),
              comma(nrow(stats_overview)),
              length(detail_gws)))
} else {
  cat('<div class="interpretation warning">No actual points data matched. Check that salary player names match stats detail names.</div>')
}
```

```{r data-quality-check, results='asis'}
# Check picked_by availability
pb_count <- sum(!is.na(data$picked_by) & as.numeric(data$picked_by) > 0, na.rm = TRUE)
cat(sprintf('<div class="interpretation">%s <strong>picked_by</strong> values detected (%.0f%% of records). %s</div>',
            comma(pb_count),
            pb_count / nrow(data) * 100,
            if (pb_count > 100) "✅ Using as primary ownership signal." else "⚠️ Falling back to salary heuristic for ownership."))
```

---

## PPG Prediction Accuracy {.tabset}

How well does our projected PPG (season average) predict this week's actual points?

```{r ppg-overall, results='asis'}
if (has_actuals) {
  ppg_data <- data %>% filter(!is.na(actual_pts), ppg > 0) %>% mutate(error = actual_pts - ppg)
  
  mae  <- round(mean(abs(ppg_data$error), na.rm = TRUE), 2)
  rmse <- round(sqrt(mean(ppg_data$error^2, na.rm = TRUE)), 2)
  bias <- round(mean(ppg_data$error, na.rm = TRUE), 2)
  r    <- round(cor(ppg_data$ppg, ppg_data$actual_pts, use = "complete.obs"), 3)
  
  cat('<div class="metric-row">')
  cat(metric_card(mae, "MAE", "neutral", "%.1f"))
  cat(metric_card(rmse, "RMSE", "neutral", "%.1f"))
  cat(metric_card(bias, "Mean Bias", if (abs(bias) < 1) "good" else "bad", "%+.1f"))
  cat(metric_card(r, "Correlation", if (r > 0.25) "good" else if (r > 0.1) "neutral" else "bad", "%.3f"))
  cat('</div>')
  
  cat(sprintf('<div class="interpretation">A PPG correlation of <strong>%.3f</strong> means our season averages explain about <strong>%.0f%%</strong> of the variance in weekly scores. Football is inherently noisy — correlations above 0.20 are useful, above 0.30 is strong.</div>', r, r^2 * 100))
}
```

### Scatter: Predicted vs Actual

```{r ppg-scatter, fig.height=7}
if (has_actuals) {
  ppg_plot <- data %>%
    filter(!is.na(actual_pts), ppg > 0) %>%
    mutate(pos = factor(pos, levels = c("GK", "DEF", "MID", "FWD")))
  
  ggplot(ppg_plot, aes(x = ppg, y = actual_pts, color = pos)) +
    geom_point(alpha = 0.3, size = 1.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = pal$muted, linewidth = 0.8) +
    geom_smooth(method = "lm", se = FALSE, linewidth = 0.8) +
    scale_color_manual(values = pos_colors) +
    facet_wrap(~pos, scales = "free") +
    labs(title = "Predicted PPG vs Actual Points",
         subtitle = "Dashed line = perfect prediction. Regression line per position.",
         x = "Predicted PPG", y = "Actual Points", color = "Position") +
    theme_backtest() +
    theme(legend.position = "none")
}
```

### By Position

```{r ppg-by-position}
if (has_actuals) {
  data %>%
    filter(!is.na(actual_pts), ppg > 0) %>%
    mutate(error = actual_pts - ppg) %>%
    group_by(Position = pos) %>%
    summarise(
      N = n(),
      `Avg Predicted` = round(mean(ppg, na.rm = TRUE), 1),
      `Avg Actual` = round(mean(actual_pts, na.rm = TRUE), 1),
      MAE = round(mean(abs(error), na.rm = TRUE), 1),
      Correlation = round(cor(ppg, actual_pts, use = "complete.obs"), 3),
      .groups = "drop"
    ) %>%
    styled_table(caption = "PPG prediction accuracy by position")
}
```

### By Gameweek

```{r ppg-by-gw, fig.height=5}
if (has_actuals && length(gws_with_pts) > 1) {
  gw_acc <- data %>%
    filter(!is.na(actual_pts), ppg > 0) %>%
    mutate(error = actual_pts - ppg) %>%
    group_by(gameweek) %>%
    summarise(
      mae = mean(abs(error), na.rm = TRUE),
      correlation = cor(ppg, actual_pts, use = "complete.obs"),
      .groups = "drop"
    )
  
  ggplot(gw_acc, aes(x = factor(gameweek))) +
    geom_col(aes(y = correlation), fill = pal$sage, alpha = 0.8) +
    geom_hline(yintercept = mean(gw_acc$correlation, na.rm = TRUE),
               linetype = "dashed", color = pal$coral, linewidth = 0.8) +
    labs(title = "PPG vs Actual Correlation by Gameweek",
         subtitle = sprintf("Dashed line = average (%.3f)", mean(gw_acc$correlation, na.rm = TRUE)),
         x = "Gameweek", y = "Correlation (r)") +
    theme_backtest()
}
```

---

## Cash Rating Analysis {.tabset}

Do Cash Locks actually outscore Hard Passes?

```{r cash-tiers, results='asis'}
if (has_actuals && "cash_rating" %in% names(data)) {
  cash_tiers <- data %>%
    filter(!is.na(actual_pts), !is.na(cash_rating)) %>%
    group_by(`Rating` = cash_rating) %>%
    summarise(
      N = n(),
      `Avg Actual` = round(mean(actual_pts, na.rm = TRUE), 1),
      `Median Actual` = round(median(actual_pts, na.rm = TRUE), 1),
      `Avg Predicted` = round(mean(ppg, na.rm = TRUE), 1),
      `Avg Cash Score` = round(mean(cash_score, na.rm = TRUE), 0),
      .groups = "drop"
    ) %>%
    arrange(desc(`Avg Cash Score`))
  
  # Monotonicity check
  actual_order <- cash_tiers$`Avg Actual`
  is_monotonic <- all(diff(actual_order) <= 0)
  
  styled_table(cash_tiers, caption = "Actual performance by Cash rating tier")
  
  cat(sprintf('<div class="interpretation%s">Tier ordering is <strong>%s</strong>. %s</div>',
              if (is_monotonic) "" else " warning",
              if (is_monotonic) "monotonic ✅" else "NOT monotonic ⚠️",
              if (is_monotonic) "Cash Locks outscore each lower tier on average — the signal works." 
              else "Some tiers are out of order — the rating needs calibration."))
}
```

### Distribution by Tier

```{r cash-violin, fig.height=6}
if (has_actuals && "cash_rating" %in% names(data)) {
  tier_order <- c("Cash Lock", "Strong Play", "Decent", "Risky", "Hard Pass")
  
  plot_data <- data %>%
    filter(!is.na(actual_pts), !is.na(cash_rating), cash_rating %in% tier_order) %>%
    mutate(cash_rating = factor(cash_rating, levels = tier_order))
  
  ggplot(plot_data, aes(x = cash_rating, y = actual_pts, fill = cash_rating)) +
    geom_violin(alpha = 0.6, color = NA) +
    geom_boxplot(width = 0.15, fill = "white", alpha = 0.8, outlier.size = 0.8) +
    scale_fill_manual(values = cash_colors) +
    labs(title = "Actual Points Distribution by Cash Rating",
         subtitle = "Box = median + IQR. Do higher tiers reliably score more?",
         x = NULL, y = "Actual Points") +
    theme_backtest() +
    theme(legend.position = "none")
}
```

### Cash Score Correlation

```{r cash-correlation, results='asis'}
if (has_actuals && "cash_score" %in% names(data)) {
  valid <- data %>% filter(!is.na(actual_pts), !is.na(cash_score))
  if (nrow(valid) > 10) {
    r <- round(cor(valid$cash_score, valid$actual_pts, use = "complete.obs"), 3)
    cat(sprintf('<div class="metric-row">'))
    cat(metric_card(r, "Cash Score vs<br>Actual Points", if (r > 0.2) "good" else if (r > 0.1) "neutral" else "bad", "%.3f"))
    cat('</div>')
  }
}
```

---

## GPP Rating Analysis {.tabset}

Do Smash Plays deliver tournament-winning ceilings?

```{r gpp-tiers}
if (has_actuals && "gpp_rating" %in% names(data)) {
  data %>%
    filter(!is.na(actual_pts), !is.na(gpp_rating)) %>%
    group_by(`Rating` = gpp_rating) %>%
    summarise(
      N = n(),
      `Avg Actual` = round(mean(actual_pts, na.rm = TRUE), 1),
      `p90 Actual` = round(quantile(actual_pts, 0.90, na.rm = TRUE), 1),
      `Max Actual` = round(max(actual_pts, na.rm = TRUE), 0),
      `Avg GPP Score` = round(mean(gpp_score, na.rm = TRUE), 0),
      .groups = "drop"
    ) %>%
    arrange(desc(`Avg GPP Score`)) %>%
    styled_table(caption = "GPP tier performance — p90 and max matter most here")
}
```

### Ceiling Outcomes by Tier

```{r gpp-ceiling-chart, fig.height=5}
if (has_actuals && "gpp_rating" %in% names(data)) {
  tier_order <- c("Smash Play", "Strong GPP", "Playable", "Meh", "Fade")
  
  plot_data <- data %>%
    filter(!is.na(actual_pts), !is.na(gpp_rating), gpp_rating %in% tier_order) %>%
    mutate(gpp_rating = factor(gpp_rating, levels = tier_order))
  
  ggplot(plot_data, aes(x = gpp_rating, y = actual_pts, fill = gpp_rating)) +
    geom_violin(alpha = 0.6, color = NA) +
    geom_boxplot(width = 0.15, fill = "white", alpha = 0.8, outlier.size = 0.8) +
    scale_fill_manual(values = gpp_colors) +
    labs(title = "Actual Points Distribution by GPP Rating",
         subtitle = "For GPP, the right tail matters — look at the upside of each tier",
         x = NULL, y = "Actual Points") +
    theme_backtest() +
    theme(legend.position = "none")
}
```

---

## Matchup Quality Signal {.tabset}

Does our matchup quality score (0–100) predict actual points?

```{r mq-quintiles}
if (has_actuals && "matchup_quality" %in% names(data)) {
  data %>%
    filter(!is.na(actual_pts), !is.na(matchup_quality)) %>%
    mutate(`MQ Quintile` = ntile(matchup_quality, 5)) %>%
    group_by(`MQ Quintile`) %>%
    summarise(
      N = n(),
      `Avg MQ` = round(mean(matchup_quality, na.rm = TRUE), 1),
      `Avg Actual Pts` = round(mean(actual_pts, na.rm = TRUE), 1),
      `Avg Predicted` = round(mean(ppg, na.rm = TRUE), 1),
      .groups = "drop"
    ) %>%
    styled_table(caption = "Quintile 5 = best matchups. Does MQ predict scoring?")
}
```

### By Position

```{r mq-by-position, fig.height=6}
if (has_actuals && "matchup_quality" %in% names(data)) {
  plot_data <- data %>%
    filter(!is.na(actual_pts), !is.na(matchup_quality), ppg > 0) %>%
    mutate(pos = factor(pos, levels = c("GK", "DEF", "MID", "FWD")))
  
  ggplot(plot_data, aes(x = matchup_quality, y = actual_pts, color = pos)) +
    geom_point(alpha = 0.2, size = 1) +
    geom_smooth(method = "lm", se = TRUE, linewidth = 0.8, alpha = 0.15) +
    scale_color_manual(values = pos_colors) +
    facet_wrap(~pos, scales = "free_y") +
    labs(title = "Matchup Quality vs Actual Points by Position",
         subtitle = "Positive slope = MQ predicts scoring. Steeper = stronger signal.",
         x = "Matchup Quality (0-100)", y = "Actual Points") +
    theme_backtest() +
    theme(legend.position = "none")
}
```

```{r mq-correlation, results='asis'}
if (has_actuals && "matchup_quality" %in% names(data)) {
  mq_cors <- data %>%
    filter(!is.na(actual_pts), !is.na(matchup_quality)) %>%
    group_by(pos) %>%
    summarise(r = round(cor(matchup_quality, actual_pts, use = "complete.obs"), 3), .groups = "drop")
  
  overall_r <- round(cor(
    data$matchup_quality[!is.na(data$actual_pts) & !is.na(data$matchup_quality)],
    data$actual_pts[!is.na(data$actual_pts) & !is.na(data$matchup_quality)],
    use = "complete.obs"), 3)
  
  cat(sprintf('<div class="interpretation">Overall MQ correlation: <strong>%.3f</strong>. By position: %s. MQ signal is typically weak at the player level (0.05–0.15) because individual variance dominates team context.</div>',
              overall_r,
              paste(sprintf("%s=%.3f", mq_cors$pos, mq_cors$r), collapse = ", ")))
}
```

---

## Floor / Ceiling Calibration {.tabset}

Are our floor (2nd worst score) and ceiling (2nd best score) estimates realistic?

```{r floor-ceiling, results='asis'}
if (has_actuals) {
  fc_data <- data %>%
    filter(!is.na(actual_pts), pts_floor > 0, pts_ceiling > 0) %>%
    mutate(
      below_floor = actual_pts < pts_floor,
      above_ceiling = actual_pts > pts_ceiling,
      in_range = actual_pts >= pts_floor & actual_pts <= pts_ceiling
    )
  
  if (nrow(fc_data) > 0) {
    pct_below  <- round(mean(fc_data$below_floor, na.rm = TRUE) * 100, 1)
    pct_range  <- round(mean(fc_data$in_range, na.rm = TRUE) * 100, 1)
    pct_above  <- round(mean(fc_data$above_ceiling, na.rm = TRUE) * 100, 1)
    
    cat('<div class="metric-row">')
    cat(metric_card(paste0(pct_below, "%"), "Below Floor", if (pct_below < 25) "good" else "bad"))
    cat(metric_card(paste0(pct_range, "%"), "In Range", if (pct_range > 40) "good" else "neutral"))
    cat(metric_card(paste0(pct_above, "%"), "Above Ceiling", if (pct_above < 35) "good" else "neutral"))
    cat('</div>')
    
    cat(sprintf('<div class="interpretation">Using trimmed 2nd worst/best from history. Target distribution: ~20%% below, ~50%% in range, ~30%% above. Current: <strong>%.0f/%.0f/%.0f</strong>. %s</div>',
                pct_below, pct_range, pct_above,
                if (pct_range > 40) "Calibration looks reasonable." else "Range is too narrow or sample too small."))
  }
}
```

### By Position

```{r floor-ceiling-pos}
if (has_actuals) {
  data %>%
    filter(!is.na(actual_pts), pts_floor > 0, pts_ceiling > 0) %>%
    mutate(
      below_floor = actual_pts < pts_floor,
      above_ceiling = actual_pts > pts_ceiling,
      in_range = actual_pts >= pts_floor & actual_pts <= pts_ceiling
    ) %>%
    group_by(Position = pos) %>%
    summarise(
      N = n(),
      `% Below Floor` = round(mean(below_floor) * 100, 1),
      `% In Range` = round(mean(in_range) * 100, 1),
      `% Above Ceil` = round(mean(above_ceiling) * 100, 1),
      `Avg Floor` = round(mean(pts_floor), 1),
      `Avg Ceil` = round(mean(pts_ceiling), 1),
      .groups = "drop"
    ) %>%
    styled_table()
}
```

---

## Ownership Estimation {.tabset}

```{r ownership-section, results='asis'}
if (has_own_actuals) {
  own_data <- data %>% filter(!is.na(actual_own_pct), !is.na(est_own_pct))
  
  own_mae  <- round(mean(abs(own_data$est_own_pct - own_data$actual_own_pct)), 2)
  own_r    <- round(cor(own_data$est_own_pct, own_data$actual_own_pct, use = "complete.obs"), 3)
  own_bias <- round(mean(own_data$est_own_pct - own_data$actual_own_pct), 2)
  
  cat('<div class="metric-row">')
  cat(metric_card(own_mae, "MAE (%pts)", if (own_mae < 5) "good" else "bad", "%.1f"))
  cat(metric_card(own_r, "Correlation", if (own_r > 0.5) "good" else if (own_r > 0.3) "neutral" else "bad", "%.3f"))
  cat(metric_card(own_bias, "Mean Bias", if (abs(own_bias) < 2) "good" else "bad", "%+.1f"))
  cat('</div>')
} else {
  cat('<div class="interpretation warning"><strong>No actual ownership data available.</strong> Add data to the ownership Google Sheet (<code>1uI0wkZVLBL9hgMDfoszuZYDx-n9P03BJPZ4vP1KAMjY</code>) in tabs named <code>week_20</code>, <code>week_21</code>, etc. with columns: <code>player, pos, team, price, own_pct, captain_pct, form, score</code>.</div>')
}
```

### Predicted vs Actual Ownership

```{r own-scatter, fig.height=6}
if (has_own_actuals) {
  own_data <- data %>% filter(!is.na(actual_own_pct), !is.na(est_own_pct))
  
  ggplot(own_data, aes(x = est_own_pct, y = actual_own_pct, color = pos)) +
    geom_point(alpha = 0.5, size = 2) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = pal$muted) +
    scale_color_manual(values = pos_colors) +
    labs(title = "Estimated vs Actual Ownership",
         subtitle = "Points on the dashed line = perfect prediction",
         x = "Estimated Ownership %", y = "Actual Ownership %") +
    theme_backtest()
}
```

### Worst Misses

```{r own-misses}
if (has_own_actuals) {
  data %>%
    filter(!is.na(actual_own_pct), !is.na(est_own_pct)) %>%
    mutate(Error = round(abs(est_own_pct - actual_own_pct), 1)) %>%
    arrange(desc(Error)) %>%
    head(15) %>%
    transmute(GW = gameweek, Player = player, Pos = pos, `Salary` = salary,
              `Est Own%` = est_own_pct, `Actual Own%` = actual_own_pct,
              Error, `Picked By` = picked_by, `Cash` = cash_rating) %>%
    styled_table(caption = "15 largest ownership estimation errors")
}
```

### Spot Check (Latest GW)

```{r own-spot-check}
# Always show top predicted ownership for manual comparison
latest_gw <- max(data$gameweek, na.rm = TRUE)

data %>%
  filter(gameweek == latest_gw) %>%
  arrange(desc(est_own_pct)) %>%
  head(20) %>%
  transmute(Player = player, Pos = pos, Team = team_normalized,
            Salary = salary, `Est Own%` = est_own_pct, `Est Cpt%` = est_captain_pct,
            `Picked By` = picked_by, Cash = cash_rating, MQ = matchup_quality) %>%
  styled_table(caption = sprintf("Top 20 predicted ownership — GW %d (compare against actual contest results)", latest_gw))
```

---

## Picked By Signal {.tabset}

How predictive is the FanTeam season-long `picked_by` statistic?

```{r picked-by, results='asis'}
if ("picked_by" %in% names(data) && has_actuals) {
  pb_data <- data %>%
    filter(!is.na(actual_pts)) %>%
    mutate(pb = as.numeric(gsub("[^0-9.]", "", as.character(picked_by)))) %>%
    filter(!is.na(pb), pb > 0)
  
  if (nrow(pb_data) > 20) {
    r_pts <- round(cor(pb_data$pb, pb_data$actual_pts, use = "complete.obs"), 3)
    r_ppg <- round(cor(pb_data$pb, pb_data$ppg, use = "complete.obs"), 3)
    
    cat('<div class="metric-row">')
    cat(metric_card(r_pts, "picked_by vs<br>Actual Points", if (r_pts > 0.15) "good" else "neutral", "%.3f"))
    cat(metric_card(r_ppg, "picked_by vs<br>Season PPG", if (r_ppg > 0.3) "good" else "neutral", "%.3f"))
    cat(metric_card(comma(nrow(pb_data)), "Players with<br>picked_by > 0", "neutral"))
    cat('</div>')
    
    cat(sprintf('<div class="interpretation"><code>picked_by</code> vs actual weekly points (r=%.3f) is the <strong>weekly signal</strong> — moderate because it is a season-long stat. picked_by vs PPG (r=%.3f) shows how well it tracks season averages — this should be high.</div>',
                r_pts, r_ppg))
  }
}
```

### Picked By Deciles vs Actual Points

```{r pb-deciles, fig.height=5}
if ("picked_by" %in% names(data) && has_actuals) {
  pb_data <- data %>%
    filter(!is.na(actual_pts)) %>%
    mutate(pb = as.numeric(gsub("[^0-9.]", "", as.character(picked_by)))) %>%
    filter(!is.na(pb), pb > 0) %>%
    mutate(pb_decile = ntile(pb, 10))
  
  if (nrow(pb_data) > 50) {
    decile_summary <- pb_data %>%
      group_by(pb_decile) %>%
      summarise(
        avg_actual = mean(actual_pts, na.rm = TRUE),
        avg_pb = mean(pb, na.rm = TRUE),
        .groups = "drop"
      )
    
    ggplot(decile_summary, aes(x = factor(pb_decile), y = avg_actual)) +
      geom_col(fill = pal$sage, alpha = 0.8) +
      geom_text(aes(label = sprintf("%.1f", avg_actual)), vjust = -0.5, size = 3.5, color = pal$primary) +
      labs(title = "Average Actual Points by Picked-By Decile",
           subtitle = "Decile 10 = most popular players. Should show upward trend.",
           x = "Picked-By Decile (1=least → 10=most popular)", y = "Avg Actual Points") +
      theme_backtest()
  }
}
```

---

## Biggest Misses {.tabset}

### Cash Locks That Bombed

Players rated as Cash Lock who scored poorly — where is the model overconfident?

```{r cash-bombs}
if (has_actuals && "cash_rating" %in% names(data)) {
  data %>%
    filter(cash_rating == "Cash Lock", !is.na(actual_pts)) %>%
    arrange(actual_pts) %>%
    head(15) %>%
    transmute(GW = gameweek, Player = player, Pos = pos, Salary = salary,
              `Cash Score` = cash_score, `Actual Pts` = actual_pts,
              PPG = ppg, MQ = matchup_quality, `Est Own%` = est_own_pct) %>%
    styled_table(caption = "Cash Locks with lowest actual points")
}
```

### Hard Passes That Exploded

Players rated as Hard Pass who massively outperformed — where are we missing value?

```{r hard-pass-heroes}
if (has_actuals && "cash_rating" %in% names(data)) {
  data %>%
    filter(cash_rating == "Hard Pass", !is.na(actual_pts)) %>%
    arrange(desc(actual_pts)) %>%
    head(15) %>%
    transmute(GW = gameweek, Player = player, Pos = pos, Salary = salary,
              `Cash Score` = cash_score, `Actual Pts` = actual_pts,
              PPG = ppg, MQ = matchup_quality, `Est Own%` = est_own_pct) %>%
    styled_table(caption = "Hard Passes with highest actual points")
}
```

### GPP Fades That Smashed

```{r gpp-fade-heroes}
if (has_actuals && "gpp_rating" %in% names(data)) {
  data %>%
    filter(gpp_rating == "Fade", !is.na(actual_pts)) %>%
    arrange(desc(actual_pts)) %>%
    head(15) %>%
    transmute(GW = gameweek, Player = player, Pos = pos, Salary = salary,
              `GPP Score` = gpp_score, `Actual Pts` = actual_pts,
              Ceiling = pts_ceiling, MQ = matchup_quality, `Est Own%` = est_own_pct) %>%
    styled_table(caption = "Fades with highest actual points — GPP upside we're missing")
}
```

---

## Data Export {.tabset}

```{r export, results='asis'}
export_cols <- c(
  "gameweek", "player", "pos", "team_normalized", "salary",
  "ppg", "pts_floor", "pts_ceiling", "sortino", "form",
  "matchup_quality", "est_own_pct", "est_captain_pct",
  "cash_score", "cash_rating", "gpp_score", "gpp_rating",
  "captain_score", "captain_rating",
  "value", "picked_by", "n_games",
  "actual_pts", "actual_own_pct", "actual_captain_pct",
  "opponent", "home_away"
)

export <- data %>%
  select(any_of(export_cols)) %>%
  arrange(gameweek, desc(cash_score))

export_path <- "data/backtest_results.csv"
dir.create(dirname(export_path), showWarnings = FALSE, recursive = TRUE)
write.csv(export, export_path, row.names = FALSE)

cat(sprintf('<div class="interpretation">Full results exported to <code>%s</code> — <strong>%s rows</strong> across <strong>%d gameweeks</strong>. Open in Excel or R for deeper analysis.</div>',
            export_path, comma(nrow(export)), length(unique(export$gameweek))))
```

---

<div style="text-align: center; color: var(--muted); padding: 2em 0; font-size: 0.85em;">
Generated `r format(Sys.time(), '%d %b %Y at %H:%M')` · Soccer Ratings Backtester v1.0
</div>
